### 2017-04-07 text/abstract.tex
minor edits throughout

\begin{abstract}

  The availability and amount of sequenced genomes have been rapidly
  growing in recent years because of the adoption of next-generation
  sequencing (NGS) technologies that emable high-throughput
  short-read generation at highly competitive cost. Since this trend
  is expected to continue in the foreseeable future, the design and
  implementation of efficient and scalable NGS bioinformatics
  algorithms are important to research and industrial applications.
  In this paper, we introduce S-Aligner---a highly scalable read
  mapper designed for the \emph{Sunway Taihu Light} supercomputer and
  its fourth-generation ShenWei many-core architecture
  (SW26010). S-Aligner employs a combination of optimization
  techniques to overcome both the memory-bound and the compute-bound
  bottlenecks in the read mapping algorithm. In order to make full use
  of the compute power of Sunway Taihu Light, our design employs three
  levels of parallelism: (1) internode parallelism using MPI based on
  a task-grid pattern, (2) intranode parallelism using
  multithreading and asynchronous data transfer to fully utilize all
  260 cores of the SW26010 many-core processor, and (3) vectorization
  to exploit the available 256-bit SIMD vector registers.  Moreover,
  we have employed asynchronous access patterns and data-sharing
  strategies during file I/O to overcome bandwidth limitations of the
  network file system. Our performance evaluation demonstrates that
  S-Aligner scales almost linearly with approximately 95\% efficiency
  for up to 13,312 nodes (concurrently harnessing more than 3 million
  compute cores). Furthermore, our implementation on a single node
  outperforms the established RazerS3 mapper running on a platform
  with eight Intel Xeon E7-8860v3 CPUs while achieving highly
  competitive alignment accuracy.

\end{abstract}