### 2017-04-08 s-aligner.tex
Added Argonne government license.

\documentclass[conference]{IEEEtran}

\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage{cases}
\usepackage{listings}
\lstset{language=C++,
    basicstyle=\ttfamily \footnotesize,
    keywordstyle=\color{black}\ttfamily,
    stringstyle=\color{black}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    morecomment=[s][\color{blue}]{/**}{**/},
    showstringspaces=false,
    showtabs=false,
    captionpos=b,
    numbers=left
}

\hyphenation{cluster computational-biology short-read-mapping}

\begin{document}

\title{S-Aligner: Ultrascalable read mapping on Sunway Taihu Light}

\author{\IEEEauthorblockN{Xiaohui Duan, Kai Xu, Yuandong Chan}
\IEEEauthorblockA{
Shandong University\\
Jinan, China\\
Email: \{sunrise.duan, xu\_kai, yuandong.c\}@mail.sdu.edu.cn}
\and
\IEEEauthorblockN{Christian Hundt, Bertil Schmidt}
\IEEEauthorblockA{Johannes Gutenberg University\\
Mainz, Germany\\
Email: \{hundt, bertil.schmidt\}@uni-mainz.de}
\and
\IEEEauthorblockN{Pavan Balaji}
\IEEEauthorblockA{Argonne National Laboratory\\
Lemont, Illinois, USA\\
Email: balaji@anl.gov}
\and
\IEEEauthorblockN{Weiguo Liu}
\IEEEauthorblockA{Shandong University\\
Jinan, China\\
Email: weiguo.liu@sdu.edu.cn\\
}
}

\maketitle

\input{text/abstract}
\IEEEpeerreviewmaketitle

\input{text/introduction}
\input{text/background}
\input{text/design}
\input{text/evaluation}
\input{text/conclusion}

\section*{Acknowledgment}
This material is based upon work supported by the U.S. Department of
Energy, Office of Science, under contract DE-AC02-06CH11357.

\bibliographystyle{plain}
\bibliography{bib/references.bib}

\vspace{20pt} \footnotesize{The submitted manuscript has been created
  by UChicago Argonne, LLC, Operator of Argonne National Laboratory
  (``Argonne'').  Argonne, a U.S. Department of Energy Office of
  Science laboratory, is operated under Contract
  No. DE-AC02-06CH11357. The U.S. Government retains for itself, and
  others acting on its behalf, a paid-up nonexclusive, irrevocable
  worldwide license in said article to reproduce, prepare derivative
  works, distribute copies to the public, and perform publicly and
  display publicly, by or on behalf of the Government.  The Department
  of Energy will provide public access to these results of federally
  sponsored research in accordance with the DOE Public Access
  Plan. http://energy.gov/downloads/doe-public-access-plan.}

\end{document}